Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-15 thesis/chapters/ch_introduction.tex
updates

@@

We dedicate this section to a specific Local Group dwarf galaxy, Leo P, as its observed properties motivate the initial conditions of the simulations developed in this dissertation. This galaxy is remarkable for being one of the lowest mass dwarf galaxies with observed, ongoing star formation. It is located at a distance of $1.62\pm 0.15$~Mpc \citep{McQuinn2015a}, and was first characterized as part of the ALFALFA survey \citep{Giovanelli2013,Rhode2013,Skillman2013,McQuinn2013,Bernstein-Cooper2014}. These studies have characterized its star formation history as low and roughly continuous over the age of the Universe (value, cite), its stellar mass ($M_{*} = 5.7 \times 10^{5}$~M$_{\odot}$, \cite{McQuinn2013}), its HI and dynamical mass ($M_{\rm HI}(r<r_{\rm HI}) = 9.5\times 10^{5}$~M$_{\odot}$, $M_{\rm dyn}(r<r_{\rm HI}) = 2.6 \times 10^{7}$~M$_{\odot}$, $r_{\rm HI} = $, \cite{Bernstein-Cooper2014}), and gas-phase abundances from an HII region (12 + log(O/H) = 7.17$\pm$0.04 \cite{Skillman2013}). Finally, we chose to focus on a Leo P like galaxy in our simulations in part to allow for comparisons to the work in \cite{McQuinn2015}, which makes an accounting of the total metals (as traced by O) contained in the ISM and stars in Leo P, estimating the fraction that must have been ejected over the lifetime of the galaxy. This is an important measurement as prior estaimtes of this number for low mass dwarf galaxies were restricted to Milky Way dSph's, which are devoid of gas and contaminated by environmental affects. As of this writing, Leo P is one of the only low mass dwarf galaxies with both gas-phase and stellar metal abundance measurements, but there is ongoing work to greatly expand this sample in the near future.

>>>>

We are motivated by using a properly formed model for galactic chemical evolution in dwarf galaxies to interpret and explain observations of Local Group dwarfs. As any good theoretical work should maintain a close connection to the observations that motivate the study in the first place in order to better make predictions for future observations, we devote some time here to summarize relevant observations of Local Group dwarfs.

\subsection{Milky Way and Local Group Dwarf Galaxies}

The Milky Way and its local environment is host to a diverse collection of dwarf galaxies. The properties of these galaxies have been nicely summarized in a recent review \citep{Tolstoy2009}, and again in \cite{McConnachie2012}. However, the number of known Local Group dwarf galaxies has increased dramatically since that time, particularly for the Milky Way, whose known satellite population has more than doubled. This is due in large part to tremendous improvements in our ability to detect galaxies on the faintest end of the luminosity function, called ultrafaint galaxies (UFDs, \cite{Willman2005}).\footnote{See \cite{Simon2019} for a recent review}. The Dark Energy Survey \citep[e.g.][]{Drlica-Wagner2015} and Hyper Supbrime-Cam \citep[e.g.][]{Greco2018} have made tremendous advances to this end, with the upcoming Large Synoptic Survey Telescope producing an even greater increase in the known population of faint galaxies in the Local Group \citep{Haynes2019,Weisz2019}.

While there is no real sharp transition where a galaxy is considered a ``dwarf" galaxy, historical definitions are based on luminosity, using the SMC and LMC as rough anchor points for the upper-end of the dwarf galaxy scale. \cite{Bullock2017} suggests a deliniation based on stellar mass. For this dissertation, we follow this definition and consider galaxies with $M_* < 10^9$~M_${\odot}$ as dwarf galaxies, and galaxies with $M_* < 10^5$~M$_{\odot}$ as UFDs.

@@

We dedicate this section to a specific Local Group dwarf galaxy, Leo P, as its observed properties motivate the initial conditions of the simulations developed in this dissertation. This galaxy is remarkable for being one of the lowest mass dwarf galaxies with observed, ongoing star formation. It is located at a distance of $1.62\pm 0.15$~Mpc \citep{McQuinn2015a}, and was first characterized as part of the ALFALFA survey \citep{Giovanelli2013,Rhode2013,Skillman2013,McQuinn2013,Bernstein-Cooper2014}. These studies have characterized its star formation history as low and roughly continuous over the age of the Universe (SFR = $4.3\times 10^{-5}$~M$_{\odot}$~yr$^{-1}$, \cite{McQuinn2015a}), its stellar mass ($M_{*} = 5.7 \times 10^{5}$~M$_{\odot}$, \cite{McQuinn2013}), its HI and dynamical mass ($M_{\rm HI}(r<r_{\rm HI}) = 9.5\times 10^{5}$~M$_{\odot}$, $M_{\rm dyn}(r<r_{\rm HI}) = 2.6 \times 10^{7}$~M$_{\odot}$, $r_{\rm HI} \sim 500$~pc, \cite{Bernstein-Cooper2014}), and gas-phase abundances from an HII region (12 + log(O/H) = 7.17$\pm$0.04 \cite{Skillman2013}). Finally, we chose to focus on a Leo P like galaxy in our simulations in part to allow for comparisons to the work in \cite{McQuinn2015}, which makes an accounting of the total metals (as traced by O) contained in the ISM and stars in Leo P, estimating the fraction that must have been ejected over the lifetime of the galaxy. This is an important measurement as prior estaimtes of this number for low mass dwarf galaxies were restricted to Milky Way dSph's, which are devoid of gas and contaminated by environmental affects. As of this writing, Leo P is one of the only low mass dwarf galaxies with both gas-phase and stellar metal abundance measurements, but there is ongoing work to greatly expand this sample in the near future.
